Charged pion emission from neutron-rich heavy ion collisions for studies on the symmetry energy

### 2020-08-05 chapters_altFormat/ch_data_analysis_2.tex
added new format

\begin{figure}[!htb]

     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{beam-Sn132.png}
         \caption{$\tin{132}{124}$ system.}
         \label{fig:beampid132}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{beam-Sn108.png}
         \caption{$\tin{108}{112}$ system.}
         \label{fig:beampid108}
     \end{subfigure}
        \caption{Beam PID achieved of the the neutron rich and poor systems, from analysis of the BigRIPs spectrometer \cite{jon} }
        \label{fig:beampidTwo}
\end{figure}

@@

\begin{figure}[!htb]

    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{clusterLR.png} 
        \caption{Top view of the tracks showing the edge effect on the left and right sides of the TPC.} 	   \label{fig:clusterLR}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{clusterTop.png} 
        \caption{Side view of the TPC showing the edge effect near the top.} \label{fig:clusterTop}
    \end{subfigure}

    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{clusterBottom.png} 
        \caption{Side view of the TPC showing the edge effect near the bottom.} \label{fig:clusterBottom}
    \end{subfigure}
\caption{Example showing the edge effect has on the cluster positions near the extreme edges of the TPC volume. }    
\label{fig:edge}
\end{figure}

@@

\begin{figure}[!htb]

     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{sideviewHighDenCut.jpg}
         \caption{Side view.}
         \label{fig:sideHigh}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{topviewHighDenCut.jpg}
         \caption{Top view.}
         \label{fig:topHigh}
     \end{subfigure}
        \label{fig:highcut}
        \caption{Views showing the high density and edge cuts.}
        \label{fig:elipsecut}
\end{figure}

@@

\begin{figure}[!htb]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{beamAngle-Sn108.png} 
        \caption{Beam angle of the $\tin{108}{124}$ system} \label{fig:beamangle108}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{beamAngle-Sn112.png} 
        \caption{Beam angle of the $\tin{112}{124}$ system} \label{fig:beamangle112}
    \end{subfigure}

    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{beamAngle-Sn124.png} 
        \caption{Beam angle of the $\tin{124}{112}$ system} \label{fig:beamangle124}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{beamAngle-Sn132.png} 
        \caption{Beam angle of the $\tin{132}{124}$ system} \label{fig:beamangle132}
    \end{subfigure}
\caption{Beam angle at the face of the target in the TPC for all 4 systems.}
\label{fig:beamangle}
\end{figure}

@@

\begin{figure}[!htb]
\centering
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{vertex-Sn108.png} 
        \caption{$\tin{108}{124}$ system.} \label{fig:vertex108}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{vertex-Sn112.png} 
        \caption{$\tin{112}{124}$ system.} \label{fig:vertex112}
    \end{subfigure}

    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{vertex-Sn124.png} 
        \caption{$\tin{124}{112}$ system.} \label{fig:vertex124}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{vertex-Sn132.png} 
        \caption{$\tin{132}{124}$ system.} \label{fig:vertex132}
    \end{subfigure}
\caption{Vertex distributions, around the target region, of all experimental runs for each secondary beam system. }
\label{fig:vertexdist}
\end{figure}

@@

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{angularCut-pim-Sn108.png}
         \caption{$\pi^-$ angular cuts in the $\tin{108}{112}$ system. $\theta_{CM} < 90$ and < $\phi$ <}
         \label{fig:pim108angle}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{angularCut-pim-Sn132.png}
         \caption{$\pi^-$ angular cuts in the $\tin{132}{124}$ system. $\theta_{CM} < 90$ and < $\phi$ <}
         \label{fig:pim132angle}
     \end{subfigure}
        \caption{Measured $\pi^-$ angular distribution for the two spherical angles $\theta$ and $\phi$.}
        \label{fig:pim}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{angularCut-pip-Sn108.png}
         \caption{$\pi^+$ angular cuts in the $\tin{108}{112}$ system. $\theta_{CM} < 90$ and < $\phi$ <}
         \label{fig:pip108angle}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{angularCut-pip-Sn132.png}
         \caption{$\pi^+$ angular cuts in the $\tin{132}{124}$ system. $\theta_{CM} < 90$ and < $\phi$ <}
         \label{fig:pip132angle}
     \end{subfigure}
        \caption{Measured $\pi^+$ angular distribution for the two spherical angles $\theta$ and $\phi$.}
        \label{fig:pip}
\end{figure}

>>>>

\begin{figure}[!htb]%
    \centering
    \subfloat[$\tin{132}{124}$ system.]{{\includegraphics[width=.46\textwidth,valign=c]{beam-Sn132.png} \label{fig:beampid132}}}%
    \qquad
    \subfloat[$\tin{108}{112}$ system.]{{\includegraphics[width=.46\textwidth,valign=c]{beam-Sn108.png} \label{fig:beampid108}}}%

	\caption{Beam PID achieved of the the neutron rich and poor systems, from analysis of the BigRIPs spectrometer \cite{jon} }
	\label{fig:beampidTwo}
\end{figure}

@@

\begin{figure}[!htb]%
    \centering
    \subfloat[Top view of the tracks showing the edge effect on the left and right sides of the TPC.]{{\includegraphics[width=.45\textwidth,valign=c]{clusterLR.png} \label{fig:clusterLR}}}%
    \qquad
    \subfloat[Side view of the TPC showing the edge effect near the top.]{{\includegraphics[width=.45\textwidth,valign=c]{clusterTop.png} \label{fig:clusterTop}}}%
    \subfloat[Side view of the TPC showing the edge effect near the bottom.]{{\includegraphics[width=.45\textwidth,valign=c]{clusterBottom.png} \label{fig:clusterBottom}}}%

  	\caption{Example showing the edge effect has on the cluster positions near the extreme edges of the TPC volume. }    
	\label{fig:edge}
\end{figure}

@@

\begin{figure}[!htb]%
    \centering
    \subfloat[Side view.]{{\includegraphics[width=.46\textwidth,valign=c]{sideviewHighDenCut.jpg} \label{fig:sideHigh}}}%
    \qquad
    \subfloat[Top view.]{{\includegraphics[width=.46\textwidth,valign=c]{topviewHighDenCut.jpg} \label{fig:topHigh}}}%
 	  \label{fig:highcut}
        \caption{Views showing the high density and edge cuts.}
        \label{fig:elipsecut}
\end{figure}

@@

\begin{figure}[!htb]%
    \centering
    \subfloat[Beam angle of the $\tin{108}{124}$ system]{{\includegraphics[width=.45\textwidth,valign=c]{beamAngle-Sn108.png} \label{fig:beamangle108}}}%
    \qquad
    \subfloat[Beam angle of the $\tin{112}{124}$ system]{{\includegraphics[width=.45\textwidth,valign=c]{beamAngle-Sn112.png} \label{fig:beamangle112}}}%

    \subfloat[Beam angle of the $\tin{124}{112}$ system]{{\includegraphics[width=.45\textwidth,valign=c]{beamAngle-Sn124.png} \label{fig:beamangle124}}}%
    \qquad
    \subfloat[Beam angle of the $\tin{132}{124}$ system]{{\includegraphics[width=.45\textwidth,valign=c]{beamAngle-Sn132.png} \label{fig:beamangle132}}}%

	\caption{Beam angle at the face of the target in the TPC for all 4 systems.}
	\label{fig:beamangle}
\end{figure}

@@

\begin{figure}[!htb]%
    \centering
    \subfloat[$\tin{108}{124}$ system.]{{\includegraphics[width=.8\textwidth,valign=c]{vertex-Sn108.png} \label{fig:vertex108}}}% 
    \qquad
    \subfloat[$\tin{112}{124}$ system.]{{\includegraphics[width=.8\textwidth,valign=c]{vertex-Sn112.png} \label{fig:vertex112}}}% 
    \\
    \subfloat[$\tin{124}{112}$ system.]{{\includegraphics[width=.8\textwidth,valign=c]{vertex-Sn124.png} \label{fig:vertex124}}}% 
    \\
    \subfloat[$\tin{132}{124}$ system.]{{\includegraphics[width=.8\textwidth,valign=c]{vertex-Sn132.png} \label{fig:vertex132}}}%

  	\caption{Vertex distributions, around the target region, of all experimental runs for each secondary beam system. }
	\label{fig:vertexdist}
\end{figure}

@@

\begin{figure}[!htb]%
    \centering
    \subfloat[$\pi^-$ angular cuts in the $\tin{108}{112}$ system. $\theta_{CM} < 90$ and < $\phi$ <]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pim-Sn108.png} \label{fig:pim108angle}}}%
    \qquad
    \subfloat[$\pi^-$ angular cuts in the $\tin{132}{124}$ system. $\theta_{CM} < 90$ and < $\phi$ <]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pim-Sn132.png} \label{fig:pim132angle}}}%

	\caption{Measured $\pi^-$ angular distribution for the two spherical angles $\theta$ and $\phi$.}
    \label{fig:pim}
\end{figure}

\begin{figure}[!htb]%
    \centering
    \subfloat[$\pi^+$ angular cuts in the $\tin{108}{112}$ system. $\theta_{CM} < 90$ and < $\phi$ <]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pip-Sn108.png} \label{fig:pip108angle}}}%
    \qquad
    \subfloat[$\pi^+$ angular cuts in the $\tin{132}{124}$ system. $\theta_{CM} < 90$ and < $\phi$ <]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pip-Sn132.png} \label{fig:pip132angle}}}%
     \caption{Measured $\pi^+$ angular distribution for the two spherical angles $\theta$ and $\phi$.}
        \label{fig:pip}
\end{figure}

### 2020-08-05 chapters_altFormat/ch_experiment.tex
added new format

\begin{figure}[!htb]
    \centering
    \begin{subfigure}{0.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{wires_open.pdf} 
        \caption{Wires open} \label{fig:wires_open}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.42\textwidth}
        \centering
        \includegraphics[width=\linewidth]{wires_closed.pdf} 
        \caption{Wires closed.} \label{fig:wires_closed}
    \end{subfigure}
\caption{Cartoon depiction of electron drift lines going through the various wire planes.}
\label{fig:wires}
\end{figure}

@@

\begin{figure}[!htb]
    \centering
    \begin{subfigure}[t]{0.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{averageTransparency.png} 
        \caption{On configurations where all wires are the same voltage.} \label{fig:ggAvgTrans}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{transparencyDeltaV.png} 
        \caption{Off configuration where adjacent wires have a voltage difference of $2 \Delta V$.} \label{fig:ggDeltaVTrans}
    \end{subfigure}

\caption{Electron transparency as measured and simulated in the on and off configurations of the gating grid.}    
\label{fig:ggTrans}
\end{figure}

@@

\begin{figure}[!htb]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{DataTrigger1.png} 
        \caption{${}^{124}$Xe primary beam trigger.} \label{fig:dataTrigger1}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{DataTrigger2.png} 
        \caption{${}^{238}$U primary beam trigger.} \label{fig:dataTrigger2}
    \end{subfigure}
\caption{Master triggers in each primary beam setting.}    
\label{fig:datatrigger}
\end{figure}

>>>>

\begin{figure}[!htb]%
    \centering
    \subfloat[Wires open]{{\includegraphics[width=.45\textwidth,valign=t]{wires_open.pdf} \label{fig:wires_open}}}%
    \qquad
    \subfloat[Wires closed]{{\includegraphics[width=.39\textwidth,valign=t]{wires_closed.pdf} \label{fig:wires_closed}}}%
    \caption{Cartoon depiction of electron drift lines going through the various wire planes.}%
    \label{fig:wires}%
\end{figure}

@@

\begin{figure}[!htb]%
    \centering
    \subfloat[On configurations where all wires are the same voltage.]{{\includegraphics[width=.45\textwidth,valign=t]{averageTransparency.png} \label{fig:ggAvgTrans}}}%
    \qquad
    \subfloat[Off configuration where adjacent wires have a voltage difference of $2 \Delta V$.]{{\includegraphics[width=.45\textwidth,valign=t]{transparencyDeltaV.png} \label{fig:ggDeltaVTrans}}}%
\caption{Electron transparency as measured and simulated in the on and off configurations of the gating grid.}    
\label{fig:ggTrans}
\end{figure}

@@

\begin{figure}[!htb]%
    \centering
    \subfloat[${}^{124}$Xe primary beam trigger.]{{\includegraphics[width=.45\textwidth,valign=t]{DataTrigger1.png} \label{fig:dataTrigger1}}}%
    \qquad
    \subfloat[${}^{238}$U primary beam trigger.]{{\includegraphics[width=.45\textwidth,valign=t]{DataTrigger2.png} \label{fig:dataTrigger2}}}%
	\caption{Master triggers in each primary beam setting.}    
	\label{fig:datatrigger}
\end{figure}

### 2020-08-05 chapters_altFormat/ch_intro.tex
added new format

\chapter{Introduction}
On August 17, 2017, a collision of two  neutron stars produced gravitational waves which were measured by the LIGO and Advanced Virgo detectors, in combination with many other astronomical observatories, marking a new era of advanced multi-messenger observations \cite{ns_2017}, dramatically increasing already very strong interesting in  questions concerning the nature of dense matter in the universe.

Mankind has been interested in the nature of the matter which composes the visible universe, and what are its \emph{constituents} or building blocks. The seminal discovery was made in 1911 by Ernest Rutherford who scattered alpha particles from a gold foil using particles emitted by an alpha source from Marie Curie. His studies led him to conclude that atoms consist of a dense positively charged atomic nucleus surrounded by a dilute electron cloud. Later,  James Chadwick  discovered that the nucleus contained neutral particles -- neutrons  -- in addition to positively charged protons as the basic constituents of nuclear matter. Remarkably, the nucleus itself accounts for 99.9\% of the mass of the atom and only \num{e-12} of the total volume, making it the densest stable form of matter on Earth.

Without an attractive force between nucleons, the Coulomb force between protons would render the nucleus unstable. It is the strong nuclear force which exerts a large attractive force  between nucleons over a very short range. This strong force is a manifestation of the fundamental interactions between the quarks which constitute each nucleon. Three quarks of two types or \emph{flavors} labeled \emph{up} and \emph{down} with nearly equal masses are needed to form each nucleon. At low energies the internal quark structure of nucleons is less important and nucleons can be thought of as fundamental particles. The strong force between nucleons is attractive only for a small distance of approximately \SIrange{1}{2}{\femto\metre} and becomes very repulsive at even shorter distances, making the nucleus very difficult to compress. It is for this reason that the distance between any pair of nucleons is a near constant value, and therefore the central density value is relatively constant over a wide range of nuclei, from the lightest to the heaviest nuclei \cite{krane}. This density is referred to as the  \emph{saturation density}, since the central density seems to saturate at a common value, namely, $\rho_0 = \SI{1.7e14}{\gram\per\centi\metre\cubed}$ or \num{0.16} nucleons \si{\per\femto\metre\cubed}.

@@

Continuing on with the Bethe-Weizacher formula, we see that there are several correction terms accounting for such effects like the surface, $a_s$, where nucleons near the surface have fewer surrounding neighbors as nucleons inside. Since the volume of the nuclei scales as $A$, radius of the nuclei would scale as $A^{1/3}$. Also there is a correction accounting for the Coulomb force, $a_c$, which is proportional to the radial extent of the nuclei $R^{-1}$; which scales as $A^{1/3}$. The asymmetry term, $a_A$, is related to the difference in energy between adding a neutron or a proton; it is typically referred to as the \emph{symmetry energy}. Although the strong force does not distinguish between protons and neutrons the overall wave function is made up of a combination of the spacial, spin, and isospin wave functions. Since nucleons are fermions,the total wave function must be anti-symmetric under particle exchange. Because of this, each spacial and spin state of a nucleon can only be occupied by one neutron and one proton, because there still is isospin asymmetry.  Because of this it is more energetically favorable to form neutron-proton pairs than to become more and more neutron or proton rich. Thus the asymmetry term above is proportional to the factor $\frac{(N - Z)^2}{A}$ which comes directly from Pauli blocking. While Eq.~\ref{eq:semiEmp} describes normal nuclear matter, at saturation density, quite well, a more general description must be made when going to higher or lower densities.

@@

Large macroscopic objects such as neutron stars are composed of mostly pure neutron matter \cite{neutronstar} with a typical size of approximately \SI{11(2)}{\kilo\metre}, though they weigh about 1.5 times the mass of our sun. Such a compact star would normally collapse inward, under its own gravitaitonal force, except for the balancing outward force arising from the strong force which exerts a large outward pressure arising from the symmetry energy at high densities.  The densities reached in a neutron star's interior is a matter of debate. In nucleonic models it is typically 2-4$\rho_o$, but in some quark models for the interior, it can range up to  9$\rho_o$ \cite{neutronstar}. To describe matter at such high densities, the energy density of the system must be described in a more general way than Eq.~\ref{eq:semiEmp}.

>>>>

\chapter{Introduction}
On August 17, 2017, a collision of two neutron stars produced gravitational waves which were measured by the LIGO and Advanced Virgo detectors, in combination with many other astronomical observatories, marking a new era of advanced multi-messenger observations \cite{ns_2017}; reinvigorating questions concerning the nature of dense matter in the universe.

Mankind has been continuously interested in the nature of the matter which composes the visible universe, and what are its constituents or building blocks. The seminal discovery was made in 1911 by Ernest Rutherford who scattered alpha particles from a gold foil using particles emitted by an alpha source from Marie Curie. His studies led him to conclude that atoms consist of a dense, positively charged atomic nucleus, surrounded by a dilute electron cloud. Later,  James Chadwick  discovered that the nucleus contained neutral particles -- neutrons  -- in addition to the known positively charged particle -- protons -- as the basic constituents of nuclear matter. Remarkably, the nucleus accounts for 99.9\% of the mass of the atom, and only \num{e-12} of the total volume, making it the densest stable form of matter on Earth.

Without an attractive force between nucleons, the Coulomb force between protons would render the nucleus unstable. It is the strong nuclear force which exerts a large attractive force  between nucleons over a very short range. This strong force is a manifestation of the fundamental interactions between the quarks which constitute each nucleon. Three quarks, of two types or \emph{flavors} labeled \emph{up} and \emph{down}, with nearly equal masses form each nucleon. At low energies the internal quark structure of nucleons is less important and nucleons can be thought of as fundamental particles. The strong force between nucleons is attractive only for a small distance of approximately \SIrange{1}{2}{\femto\metre} and becomes very repulsive at even shorter distances, making the nucleus very difficult to compress. It is for this reason that the distance between any pair of nucleons is a near constant value, and therefore the central density value is relatively constant over a wide range of nuclei, from the lightest to the heaviest nuclei \cite{krane}. This density is referred to as the  \emph{saturation density}, since the central density seems to saturate at a common value, namely, $\rho_0 = \SI{1.7e14}{\gram\per\centi\metre\cubed}$ or \num{0.16} nucleons \si{\per\femto\metre\cubed}.

@@

Continuing on with the Bethe-Weizacher formula, we see that there are several correction terms accounting for such effects like the surface, $a_s$, where nucleons near the surface have fewer surrounding neighbors as nucleons inside. Since the volume of the nuclei scales as $A$, radius of the nuclei would scale as $A^{1/3}$. Also there is a correction accounting for the Coulomb force, $a_c$, which is proportional to the radial extent of the nuclei $R^{-1}$; which scales as $A^{1/3}$. The asymmetry term, $a_A$, is related to the difference in energy between adding a neutron or a proton; it is typically referred to as the \emph{symmetry energy}. Although the strong force does not distinguish between protons and neutrons the overall wave function is made up of a combination of the spacial, spin, and isospin wave functions. Since nucleons are fermions,the total wave function must be anti-symmetric under particle exchange. Because of this, each spacial and spin state of a nucleon can only be occupied by one neutron and one proton, because there still is isospin asymmetry.  Because of this it is more energetically favorable to form neutron-proton pairs than to become more and more neutron or proton rich. Thus the asymmetry term above is proportional to the factor $(N - Z)^2/A$ which comes directly from Pauli blocking. While Eq.~\ref{eq:semiEmp} describes normal nuclear matter, at saturation density, quite well, a more general description must be made when going to higher or lower densities.

@@

Large macroscopic objects such as neutron stars are composed of mostly pure neutron matter \cite{neutronstar} with a typical size of approximately $11^{+2}_{-2}$ \si{\kilo\metre}, though they weigh about 1.5 times the mass of our sun. Such a compact star would normally collapse inward, under its own gravitaitonal force, except for the balancing outward force arising from the strong force which exerts a large outward pressure arising from the symmetry energy at high densities.  The densities reached in a neutron star's interior is a matter of debate. In nucleonic models it is typically 2-4$\rho_o$, but in some quark models for the interior, it can range up to  9$\rho_o$ \cite{neutronstar}. To describe matter at such high densities, the energy density of the system must be described in a more general way than Eq.~\ref{eq:semiEmp}.

### 2020-08-05 chapters_altFormat/ch_summary.tex
added new format

The need for precise experimental constraints on the symmetry energy at high densities is paramount to understanding exotic, dense matter, such as neutron stars. Direct observation of these objects provides information about the total equation of state, which is primarily sensitive to the pressure in neutron matter. Only by laboratory measurements can one control the asymmetry of matter and isolate its dependence on the isospin asymmetry $\delta = \frac{\rho_n - \rho_p}{\rho}$. This information is critical to understanding the symmetry energy and associated isovector mean field potentials. These mean field potential contribute strongly to the chemical potentials, which control the phase transitions and the internal structure within stars. Heavy ion collisions provide the only laboratory environment to study these issues at the suprasaturation densities relevant to the neutron star interior.

@@

  This marks the first time pions have been accurately measured in the sub-threshold region. We observed significant a deviation from the  n\"aive expectation of the simple $\Delta$ resonance model, where $\pi^-/\pi^+ ~ (N/Z)^2$, as discussed in Section~\ref{sec:pionObs}. Such a deviation is also predicted by theory, so this explanation should be retired in favor of something more quantitative and informative. Obtaining a better explanation from current calculations is difficult because nearly all pion production calculations are incomplete and each model has a mixture of strengths and inadequacies that are unique to each model. Thus, it is difficult to get the theorists to agree about the consequences of some of the model assumptions that they employ; and consequently a constraint on the density dependence of the symmetry energy.

>>>>

The need for precise experimental constraints on the symmetry energy at high densities is paramount to understanding exotic, dense matter, such as neutron stars. Direct observation of these objects provides information about the total equation of state, which is primarily sensitive to the pressure in neutron matter. Only by laboratory measurements can one control the asymmetry of matter and isolate its dependence on the isospin asymmetry $\delta =(\rho_n - \rho_p)/\rho$. This information is critical to understanding the symmetry energy and associated isovector mean field potentials. These mean field potential contribute strongly to the chemical potentials, which control the phase transitions and the internal structure within stars. Heavy ion collisions provide the only laboratory environment to study these issues at the suprasaturation densities relevant to the neutron star interior.

@@

  This marks the first time pions have been accurately measured in the sub-threshold region. We observed significant a deviation from the  n\"aive expectation of the simple $\Delta$ resonance model, where $\pi^-/\pi^+ \sim (N/Z)^2$, as discussed in Section~\ref{sec:pionObs}. Such a deviation is also predicted by theory, so this explanation should be retired in favor of something more quantitative and informative. Obtaining a better explanation from current calculations is difficult because nearly all pion production calculations are incomplete and each model has a mixture of strengths and inadequacies that are unique to each model. Thus, it is difficult to get the theorists to agree about the consequences of some of the model assumptions that they employ; and consequently a constraint on the density dependence of the symmetry energy.
